### 2020-06-21 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
OptLet перевод в tex

\section{Problem statement}
\label{sec:2}

The multicriteria (or vector) optimization (MCO) problem can be formulated as:
\begin{equation}\label{eq:1}
f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min, \; y \in D,
\end{equation}
where $f(y) = (f1(y),f2(y),\dots,fs(y))$  is the vector efficiency criterion, $y = (y_1,y_2,\dots,y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved. The search domain $D$ defines the set of possible parameter values and commonly has a shape of an $N$-dimensional hypercube
\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
for the given boundary vectors $a$ and $b$.

@@

In the problem (\ref{eq:1}), the efficiency criteria are usually contradictory and there is no decision $y \in D$ that would provide the best (smallest) values for all criteria at the same time, i.e.,
\begin{equation}\label{eq:4}
\nexists y^*\in D: f_i(y^*) = \min_{y \in D} {f_i (y)} , 1 \leq i \leq s.
\end{equation}
If the relation (\ref{eq:4}) is valid in the solution of the MCO problem, then such decisions $\widetilde{y}^* \in D$ can be determined, at which the values of particular criteria cannot be improved without worsening the values of other criteria $f_i(y)$, $1 \leq i \leq s$,
\begin{equation}\label{eq:5}
\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*), 1 \leq i \leq s), (\exists j: f_j(y') \leq f_j(\widetilde{y}^*)).
\end{equation}

@@

\section{The Approach: Key Ideas and Methods}
\label{sec:3}

The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with non-convex constraints. Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality. The proposed approach for solving such problems is based on three key ideas:
\begin{itemize}
  \item reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
  \item using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
  \item repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
\end{itemize}

\subsection{Dimension reduction of multidimensional multicriteria optimization problems}

The search for numerical estimates of globally optimal points involves the construction of grids covering the search domain $D$ (see, e.g., \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}). This key feature of global optimization problems leads to the ``curse of dimensionality'', where the computational complexity of constructing coverage increases exponentially as the dimensions of optimization problems are enlarged. Significantly reduced complexity can be achieved by using heterogeneous grids that are denser only in the vicinity of globally optimal points. Such heterogeneous coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next iterations, the global search is performed taking into account all the available search information (points of previous search iterations and values of the optimized function at these points), that is
\begin{equation}\label{eq:9q}
y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
\end{equation}
where $k$ is the number of the global search iterations being performed, $G_k$ is the decision rule of the applied global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ step of the global search.

Adaptive selection of points $y^{k+1} \in D$ at the iterations of the global search involves the analysis of volumetric multidimensional search information $\omega_k$, $k \gg 1$, which determines the high computational complexity of the decision rules $G_k$, $k=1,2,\dots$, of global search algorithms. This computational complexity can be significantly simplified by reducing the optimization problems to be solved using Peano \textit{space-filling curves} or \textit{evolvents} $y(x)$, that uniquely and continuously map the interval $[0,1]$ to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}). As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:5}) is reduced to a one-dimensional problem:
\begin{equation}\label{eq:10q}
f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots , f_s(y(x)) \to min_{leq},  x \in [0,1].
\end{equation}

It should be noted that the one-dimensional functions obtained as a result of the reduction satisfy the uniform H\"older condition (see \cite{c17,c18}), that is
\begin{equation}\label{eq:11}
|f_i (y(x'))-f_i (y(x''))| \leq H_i \|x'-x''\|, \; x',x''\in [0,1],\; 1 \leq i \leq s,
\end{equation}
where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$, $1 \leq i \leq m$, $L_i$ are the Lipschitz constants from (\ref{eq:3}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).

Using dimension reduction the search information $\omega_k$ can be converted from (\ref{eq:9}) to the form
\begin{equation}\label{eq:12}
A_k=\{(x_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
\end{equation}
in which, in addition to the reduced representation, the data is arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$, $1 \leq i \leq k$, that is
\begin{equation}
x_1 < x_2 < \dots < x_k	
\end{equation}
for a more efficient execution of global search algorithms.

Dimension reduction using Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}. Numerical methods for constructing approximations of Peano curves (\textit{evolvents}) with a given accuracy are considered in \cite{c17,c18}.

\subsection{An efficient method for solving global optimization problems with 
non-convex constraints}

The problems solved in the framework of the computational scheme (\ref{eq:7}) are one-dimensional global optimization problems with non-convex constraints
\begin{equation}\label{eq:13}
\min{\{\varphi(x):g_i(x)\leq 0, \; 1 \leq i \leq m,\; x\in \Delta=[0,1]\}}
\end{equation}
(the notation $g_{m+1}(x) = \varphi(x)$ will also be used later).

The presence of non-convex constraints significantly complicates the complexity of solving global optimization problems; the resulting solutions must belong to a feasible domain
\begin{equation}\label{eq:14}
\Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0,   1 \leq i \leq m \}.
\end{equation}

To reduce the complexity of solving optimization problems with constraints, simpler cases are often highlighted; for example, problems with linear or quadratic constraints are selected. Various methods of approximating complex constraints with the aid of simpler forms (linear, convex, etc.) are widely used. However, the penalty function method is most commonly used, when calculating in an unfeasible region $\Delta \setminus \Delta_g$, some ``penalty'' is added to the values of the minimized function. But the penalty function method often involves repeatedly solving the optimization problem to determine a sufficient value of the penalty coefficient. More complete information on methods for solving optimization problems with constraints can be obtained, for example, in monographs \cite{c18,c26,c27,c28,c29}.

The proposed approach is based on the original method of separate accounting for constraints developed in the framework of the information-statistical theory of multiextremal optimization \cite{c18}. The key idea of the approach consists in constructing a certain integral objective function without constraints, the solution of which also provides a solution to the initial constrained optimization problem (\ref{eq:13}); a more detailed presentation of this method is given below.

We apply a sequential scheme to obtain the constraint values $g_i(y(x))$, $1 \leq i \leq m$, and the objective function $\varphi(x)$, according to which the constraint values are calculated strictly in the order of their numbers, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints in which the values were calculated will be indicated below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:15}
\nu(x) : g_\nu(x)>0,\; g_j(x)\leq 0,\; 1 \leq j\leq \nu - 1,\; 1\leq \nu = \nu(x) \leq m
\end{equation}
 (when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index concept allows us to determine the integrated function for the problem (\ref{eq:13}) (see Fig. \ref{fig:1})
\begin{equation}\label{eq:16}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (below, this procedure is called a \textit{trial}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An integrated function for the problem (\ref{eq:13}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:16}), b) the function $\Phi(x)$ from (\ref{eq:17})}
  \label{fig:1}
\end{figure}

The function $F(x)$ consists of separate fragments of the constraints $g_i(x)$, $1 \leq i \leq m$, and the objective function $\varphi(x)$. For the homogeneity of the components, the function $F(x)$ is then converted to the form 
\begin{equation}\label{eq:17}
\Phi (x) = 
 \begin{cases}
   g_\nu(x) / H_\nu, & \nu < M, \\
   (g_M(x)-z^*_M) / H_M, & \nu = M,
 \end{cases}
\end{equation}
where $M$ is the constraint number for which there are no feasible points (the maximum value of the index is $\nu(x)$), $z^*_M$ is the value of the minimum constraint violation $g_M(x)$\footnote{If $M=m+1$, then $z_M^*$ is the minimum value of the function $\varphi(x)$.}, and $H_\nu$, $1\leq \nu \leq M$, are the H\"older constants from (\ref{eq:11}). As a rule, these values are a priori unknown, but when performing calculations, instead of these values, their adaptive estimates obtained in the process of calculations can be applied. For this, the search information $A_k$ from (\ref{eq:12}) should be expanded with data
\begin{equation}\label{eq:18}
A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f( y(x_i))): 1 \leq i \leq k \},
\end{equation}
where $z_i= \Phi(x_i)$, $1 \leq i \leq k$. Then estimates of the required values can be determined as follows
\begin{equation}\label{eq:19}
M^k= \max{\{\nu=\nu(x_i), 1\leq i \leq k \}},
\end{equation}

\begin{equation}\label{eq:20}
z^k_\nu = 
 \begin{cases}
   0, & \nu < M, \\
   \min{\{g_\nu(x_i):\nu=\nu(x_i), 1\leq i \leq k \}}, & \nu = M,
 \end{cases}
\end{equation}

\begin{equation}\label{eq:21}
h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j},\; \nu_i=\nu_j=l,\; i>j, \;1\leq l\leq M^k\}}.
\end{equation}

If $h_l^k$ in (\ref{eq:21}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken. For the initial values of these quantities, $M^0 = H_l^0=1$, $z_l^0=0$, $1 \leq i \leq m+1$, can be taken.

The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method.} (AGCS) is as follows \cite{c18}.

The first trial is carried out at an arbitrary internal point $x^1\in (0,1)$. The calculation results are constantly added to the set $A_k$ from (\ref{eq:12}); in addition, into $A_k$ are also added the points $x_0=0$ and $x_{k+1}=1$ (the values $z_0$ and $z_{k+1}$ are not defined and not used) for the convenience of subsequent formulations. The choice of the point $x^{k+1}$, $k\geq 1$, of any next trial is determined by the following rules.

\textit{Rule 1.} For each interval $(x_{i-1},x_i)$, $1 \leq i \leq k+1$, in the set $A_k$ calculate the characteristic $R(i)$, where
\begin{equation}\label{eq:22}
\begin{matrix}
R(i) = 
 \begin{cases}
   \rho_i + \frac{(z_i-z_{i-1})^2}{r^2_\nu h^2_\nu \rho_i} - 2 \frac{z_i+z_{i-1}-2z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})=\nu(x_i), \\
   2 \rho_i - 4\frac{z_i-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})<\nu(x_i), \\
2 \rho_i - 4\frac{z_{i-1}-z^*_\nu}{r_\nu h_\nu}, & \nu=\nu(x_{i-1})>\nu(x_i), 
 \end{cases} \\
\rho_j = \sqrt[N]{x_i-x_{i-1}}, 1\leq j\leq k+1.
\end{matrix}
\end{equation}

The constants $r_\nu>1$, $1\leq \nu \leq m+1$, are the parameters of the algorithm. The products $r_\nu h_\nu$ are used as an estimate of the H\"older constants $H_\nu$, $1\leq \nu \leq m+1$ from (\ref{eq:11}).

\textit{Rule 2.} Determine the interval $(x_{t-1},x_t)$ that corresponds to the maximum characteristic
\begin{equation}\label{eq:23}
R(t) = \max{\{R(i):1 \leq i \leq k+1\}}.
\end{equation}

\textit{Rule 3.} Execute a new trial at the point $x^{k+1} \in (x_{t-1},x_t)$, determined in accordance with the expression
\begin{equation}\label{eq:24}
x^{k+1} = 
 \begin{cases}
   \frac{x_t+x_{t-1}}{2}, & \nu(x_{t-1}) \neq \nu(x_t), \\
   \frac{x_t+x_{t-1}}{2} + sign(z_t - z_{t-1}) \frac{1}{2r_\nu}[ \frac{|z_t - z_{t-1}|}{h_\nu}^N ], & \nu = \nu(x_{t-1}) = \nu(x_t).
 \end{cases}
\end{equation}

Iterations of the algorithm are terminated if the stop condition is satisfied
\begin{equation}\label{eq:25}
\rho_t\leq \varepsilon,
\end{equation}
where $t$ from (\ref{eq:23}) and $\varepsilon > 0$ is the given accuracy.

Results of using the AGCS algorithm to solve the test problem from Fig.~\ref{fig:1} with the parameters $r_1=r_2=r_3=2$ and $\varepsilon=10^{-5}$ are shown in Fig. \ref{fig:2}.  The coordinates of the trial points performed by the algorithm in the process of solving the problem are marked in Fig. \ref{fig:1} by three rows of vertical strokes. The strokes in the upper row correspond to points with a unit index, the second row to points whose indices equal to 2; the points marked with strokes in the bottom row belong to the feasible domain. The coordinates of trials performed at close points are marked with a dark rectangle. In total, the value of the first constraint was computed 147 times, the value of the second constraint -- 84 times, and the value of the minimized function was calculated only 35 times.

@@

\begin{theorem}
If the AGCS algorithm is used to solve problem (\ref{eq:13}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j$, $1 \leq j \leq m+1$,  satisfy the H\"older condition of (\ref{eq:11}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the quantities $h_\nu$ from (\ref{eq:21}), starting from a certain step, the inequalities are valid
\begin{equation*}
r_\nu h_\nu>2H_\nu, 1\leq \nu \leq m+1,
\end{equation*}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$, generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:13}) with $\varepsilon=0$ in the stop condition (\ref{eq:25}).
\end{theorem}

\subsection{Acceleration of computations based on the re-use of search information}

As noted earlier, the MCOlex problem at each stage of the successive concessions method (\ref{eq:7}) is a global optimization problem with non-convex constraints (\ref{eq:13}). Generally, the solution to such problems must start at the beginning, and multiple solutions of such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:18}) allows us to bring the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:7}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$,  of the initial problem (\ref{eq:1}) \cite{c30,c31}. This re-use of search information, to find the next efficient decision, reduces the amount of computations needed to solve each next optimization problem to just a few iterations; see Section \ref{sec:4} for the results of computational experiments.

Moreover, using search information $A_k$ allows to reduce the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:7}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:7}), namely 
\begin{equation}\label{eq:26}
P_{\delta lex} (f,\delta,D)=Arg \min_{y\in D}{f_s(y)}, f_j(y)\leq f_j^*+\delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, can be used the estimates $z_j^k$, $1 \leq j < s$, from (\ref{eq:20}) obtained from the search information $A_k$,  and for the constants $\delta_j$, $1 \leq j<s$, we can use the values normalized to the range of the constraints' values, that is
\begin{equation}\label{eq:27}
	\begin{matrix}
	\delta_j=z_j^k+\theta_j (g_j^k-z_j^k ),\; 0\leq \theta_j\leq 1, \; 1 \leq j < s,\\
	g_j^k= \max{\{ g_j (x_i ) : \nu(x_i)=j,\; 1\leq i\leq k \}}.
	\end{matrix}
\end{equation}

The applicability of such a one-step scheme (\ref{eq:26})--(\ref{eq:27}) instead of the scheme (\ref{eq:7}) is computationally evaluated in Section \ref{sec:4}.

>>>>

\section{Problem Statement}
\label{sec:2}

The multicriteria (or vector) optimization (MCO) problem can be formulated as:
\begin{equation}\label{eq:1}
f(y) = (f_1(y), f_2(y), \dots , f_s(y)) \to min, \; y \in D,
\end{equation}
where $f(y) = (f_1(y),f_2(y),\dots,f_s(y))$  is the vector efficiency criterion, $y = (y_1,y_2,\dots,y_N)$ is the vector of variable parameters, and $N$ is the dimension of the multicriteria optimization problem to be solved. The search domain $D$ defines the set of possible parameter values and commonly has a shape of an $N$-dimensional hypercube
\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
for the given boundary vectors $a$ and $b$.

@@

In the problem (\ref{eq:1}), the efficiency criteria are usually contradictory and there is no decision $y \in D$ that would provide the best (smallest) values for all criteria at the same time, i.e.,
\begin{equation}\label{eq:4}
\nexists y^*\in D: f_i(y^*) = \min_{y \in D} {f_i (y)} , 1 \leq i \leq s.
\end{equation}
If the relation (\ref{eq:4}) is valid in the solution of the MCO problem, then such decisions $\widetilde{y}^* \in D$ can be determined, at which the values of particular criteria cannot be improved without worsening the values of other criteria $f_i(y)$, $1 \leq i \leq s$,
\begin{equation}\label{eq:5}
\nexists y'\in D: (f_i(y') \leq f_i(\widetilde{y}^*), 1 \leq i \leq s), (\exists j: f_j(y') < f_j(\widetilde{y}^*)).
\end{equation}

@@

\section{The Proposed Approach: Key Ideas and Methods}
\label{sec:3}

The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions for global optimization (GO) problems with nonlinear constraints. Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'' -- computational complexity increases exponentially with increasing dimensionality.

For many years, global optimization has been the subject of many scientific studies. The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}. Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following:

\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function in the point of the global minimum should be compared with the values of the function in all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points,
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on eof the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function in these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

Given the requirements listed above, the proposed approach for solving MMCOlex problems is based on three key ideas:

\begin{itemize}
	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
	\item Intensive search information obtained in the process of computations for solving multiple global optimization problems.
\end{itemize}

\subsection{Dimension Reduction of Multidimensional Multicriteria Optimization Problems}

The nonuniform coverage of the search domain $D$ can be constructed adaptively in the following manner: when choosing points for the next optimization iterations, the global search is performed taking into account all the available search information (the points of previous search iterations and the values of the optimized function in these points), that is
\begin{equation}\label{eq:11}
y^{k+1}=G_k(\omega_k), \omega_k=\{(y^i,f^i=f(y^i)): 1 \leq i \leq k \},
\end{equation}
where $k$ is the number of the global search iterations performed, $G_k$ is the decisive rule of the global search algorithm, according to which the point $y^{k+1}$ of the next iteration is selected, and $\omega_k$ is the search information available at the $k$ iteration of the global search. The computational complexity of the decisive rule can be significantly simplified by reducing the optimization problems by the Peano space-filling curves or evolvents $y(x)$ that uniquely and continuously map the interval [0,1] to an $N$-dimensional hypercube $D$ (see, e.g., \cite{c17,c18,c23}). As a result of this reduction, the initial multidimensional multicriteria optimization problem (\ref{eq:7}) is reduced to a one-dimensional problem:
\begin{equation}\label{eq:12}
f(y(x)) = (f_1(y(x)), f_2(y(x)), \dots, f_s(y(x))) \to min_{lex},  x \in [0,1].
\end{equation}

It should be noted that the one-dimensional functions obtained as a result of the reduction satisfy the uniform H\"older condition (see \cite{c17,c18}), that is
\begin{equation}\label{eq:13}
|f_i (y(x'))-f_i (y(x''))| \leq H_i |x'-x''|^{1/N}, x',x''\in [0,1], 1 \leq i \leq s,
\end{equation}
where the constants $H_i$ are determined by the ratio $H_i=2L_i\sqrt{N+3}$,  $1 \leq i \leq m$, the values $L_i$ are the Lipschitz constants from (\ref{eq:2}), and $N$ is the dimension of the optimization problem (\ref{eq:1}).

Using dimension reduction, the search information $\omega_k$ can be converted from (\ref{eq:11}) to 
\begin{equation}\label{eq:14}
A_k=\{(x_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
\end{equation}
where, in addition to the reduced representation, the data is arranged in ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$, $1 \leq i \leq k$, that is
\begin{equation} \label{eq:15}
x_1 < x_2 < \dots < x_k	
\end{equation}
for more efficient execution of global search algorithms.

Dimension reduction by the Peano curves is widely used in the development of efficient global search algorithms in the framework of the information-statistical theory of multiextremal optimization \cite{c17,c18,c23,c26,c27}. Numerical methods for constructing approximations of the Peano curves (evolvents) with a given accuracy are considered in \cite{c17,c18}.

\subsection{An Efficient Method for Solving Global Optimization Problems with Nonlinear Constraints}

The problems solved in the framework of the computational scheme (\ref{eq:9}) are one-dimensional global optimization problems with nonlinear constraints
\begin{equation}\label{eq:16}
\min{\{\varphi(x):g_i(x)\leq 0, 1 \leq i \leq m, x\in \Delta=[0,1]\}}
\end{equation}
(the expression $g_{m+1}(x) = \varphi(x)$ will also be used later).

The presence of nonlinear constraints significantly increasesthe complexity of global optimization problems; the resulting solutions must belong to a feasible domain
\begin{equation}\label{eq:17}
\Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0, 1 \leq i \leq m \}.
\end{equation}

To reduce the complexity of solving optimization problems with constraints, simpler cases are often highlighted; for example, problems with linear or quadratic constraints are selected. Various methods of approximating complex constraints with the aid of simpler forms (linear, convex, etc.) are widely used. However, the most commonly used method is the penalty function method -- when calculating in an unfeasible region $\Delta \setminus \Delta_g$, certain ``penalties'' are added to the values of the minimized function. However, the penalty function method often involves the repeated solution of the optimization problem to determine a sufficient value of the penalty coefficient. More complete information on methods for solving optimization problems with constraints is presented, for example, in monographs \cite{c28,c29}.

The proposed approach is based on the original method of separate accounting for constraints developed in the framework of the information-statistical theory of multiextremal optimization \cite{c18}. The key idea of the approach is to construct an integral objective function without constraints, the solution of which also provides a solution to the initial constrained optimization problem (\ref{eq:16}); a more detailed presentation of this method is given below.

To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:18}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
\end{equation}
(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1})
\begin{equation}\label{eq:19}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) the function $F(x)$ from (\ref{eq:19}), b) the function $\Phi(x)$ from (\ref{eq:20})}
  \label{fig:1}
\end{figure}

The function $F(x)$ consists of separate fragments of the constraints ${g_i(x) \leq 0}$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. For the homogeneity of the components, the function $F(x)$ is converted to the form 
\begin{equation}\label{eq:20}
\Phi (x) = 
 \begin{cases}
   g_\nu(x) / H_\nu, & \nu < M, \\
   (g_M(x)-z^*_M) / H_M, & \nu = M,
 \end{cases}
\end{equation}
where $M$ is the constraint number, for which there are no feasible points (the maximum value of the index $\nu(x)$), $z^*_M$ is the value of the minimum constraint violation $g_M(x)$\footnote{If $M=m+1$, then $z_M^*$ is the minimum value of the function $\varphi(x)$.}, and $H_\nu$, $1\leq \nu \leq M$, are the H\"older constants from (\ref{eq:13}). As a rule, these values are a priori unknown, but when performing calculations, instead of these values, their adaptive estimates can be applied. For this, the search information $A_k$ from (\ref{eq:14}) should be expanded with data
\begin{equation}\label{eq:21}
A_k=\{(x_i, z_i, \nu_i, y(x_i), f_i=f(y(x_i))): 1 \leq i \leq k \},
\end{equation}
where $z_i= \Phi(x_i)$, $1 \leq i \leq k$. Then the estimates of the required values can be determined as follows
\begin{equation}\label{eq:22}
M^k = \max{\{\nu=\nu(x_i), 1\leq i \leq k \}},
\end{equation}

\begin{equation}\label{eq:23}
z^k_\nu = 
 \begin{cases}
   0, & \nu < M^k, \\
   \min{\{g_\nu(x_i):\nu=\nu(x_i), 1\leq i \leq k \}}, & \nu = M^k,
 \end{cases}
\end{equation}

\begin{equation}\label{eq:24}
h^k_l=\max{\{|z_i-z_j|/\sqrt[N]{x_i-x_j}, \nu_i=\nu_j=l, i>j, 1\leq l\leq M^k\}}.
\end{equation}

If $h_l^k$ in (\ref{eq:24}) turns out to be indefinite or equal to zero, then $h_l^k=1$ is taken. For the initial values of these parameters, $M^0 = H_l^0 = 1$, $z_l^0 = 0$, $1 \leq i \leq m+1$, can be taken.

In the framework of the proposed approach to minimize the function $\Phi(x)$, the global search algorithm for multiextremal optimization problems with nonlinear constraints\footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is applied. The general computational scheme of this method can be presented as follows \cite{c18}.

The first trial is carried out at an arbitrary internal point $x^1\in(0,1)$. The choice of the point $x^{k+1}$, $k \geq 1$ of any next trial is determined by the following rules

\textit{Rule 1.} For each interval $(x_{i-1},x_i)$, $1 < i \leq k$, in the set $A_k$ calculate the characteristic

\begin{equation}\label{eq:25}
R(i)=\mathcal{R} (i,A_k),
\end{equation}
where the expression $\mathcal{R}(i,A_k)$ is set by AGCS.

\textit{Rule 2.} Determine the interval $(x_{t-1},x_t)$ that corresponds to the maximum characteristic
\begin{equation}\label{eq:26}
R(t) = \max{\{R(i):1 < i \leq k\}}.
\end{equation}

\textit{Rule 3.} Execute a new trial at the point $x^{k+1} \in (x_{t-1},x_t)$, determined in accordance with the expression
\begin{equation}\label{eq:27}
x^{k+1} = \mathcal{X}(x_{t-1},x_t)
\end{equation}
where the expression $\mathcal{X}(x_{t-1},x_t)$ is also set by AGCS.

Iterations of the algorithm are terminated if the stop criterion is satisfied

\begin{equation}\label{eq:28}
\rho_t\leq \varepsilon,
\end{equation}
where $t$ from (\ref{eq:26}), $\rho_t = \sqrt[N]{x_t-x_{t-1}}$ and $\varepsilon > 0$ is the given accuracy.

Results of the application of the AGCS algorithm to solve the test problem from Fig.~\ref{fig:1} with the accuracy $\varepsilon=10^{-5}$ are shown in Fig.~\ref{fig:2}. The coordinates of the trial points performed by the algorithm in the process of solving the problem are marked in Fig.~\ref{fig:2} by three rows of vertical strokes. The strokes in the upper row correspond to points with a unit index, the second row to points, the indices of which equal to $\nu=2$; the points marked with strokes in the bottom row belong to the feasible domain. The coordinates of trials performed at close points are marked with a dark rectangle. In total, the value of the first constraint was computed 147 times, the value of the second constraint -- 84 times, and the value of the minimized function was calculated only 35 times.

@@

\begin{theorem}
If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j$, $1 \leq j \leq m+1$, $[0,1]$ satisfy the H\"older condition of (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the values $h_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$,are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
\begin{equation} \label{eq:29}
r_\nu h_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
\end{equation}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
\end{theorem}

\subsection{Acceleration of Computations Based on the Reuse of Search Information}

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of computational experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}

In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.
